Describing the C-Space of Asynchronous Periodic Task Systems using Definitive Idle Times

### 2013-07-29 dit-paper_ieee_old.tex

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%
\documentclass[conference]{IEEEtran}
\usepackage{cite}
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
\fi
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{url}

% added
\usepackage{amssymb}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{stmaryrd}

@@

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Describing the C-Space of Asynchronous Periodic Task Systems using
Definitive Idle Times}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Thomas Chapeaux\\ and Paul Rodriguez}
\IEEEauthorblockA{Universit\'{e} Libre de Bruxelles\\
Email: tchapeau@ulb.ac.be, paurodri@ulb.ac.be}
\and
\IEEEauthorblockN{Laurent George\\ and Jean-Fran\c{c}ois Hermant}
\IEEEauthorblockA{ECE Paris\\
Email}
}
% make the title area
\maketitle

\begin{abstract}
%\boldmath
Sensitivity analysis for real-time systems asks for efficient methods to determine feasibility
after small changes of execution times, such as when porting a system to a different platform.
The C-space, presented in \cite{bini2004schedulability,george2009characterization}, describes the
set of execution times for which a system is feasible. This paper extends previous works to
generalize the description of the C-space for Earliest Deadline First scheduling to asynchronous
systems and non-pairwise coprime periods. The feasibility gain offered by offsets is then
expressed as the volume ratio between the C-space of an asynchronous system and its corresponding
synchronous system.
\end{abstract}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

	Abstract models for real-time systems often require the determination of the
	(worst-case) computation time of a job (WCET), a value highly dependent on the
	platform on which the application is deployed.

	The \emph{C-space} of a real-time system is, intuitively,
	the set of WCET values for which this system would be feasible. A
	precise and efficient description of its C-space would allow to quickly decide
	whether or not a system is feasible on a given platform.

@@

		We consider a discrete timeline and tasks systems $\tau$ comprised of periodic concrete
		tasks with constrained deadlines on uniprocessor platforms. Each task
		$\tau_i$ is represented by a tuple $(O_i, C_i, D_i, T_i)$ where $O_i$ is the offset value,
		$C_i$ the WCET, $D_i$ the relative deadline and $T_i$ the period.
		The constrained deadline property implies that $D_i \leq T_i \; \forall i$. In
		the following, we differentiate between \emph{synchronous systems} (where $O_i
		= 0 \; \forall i$) and \emph{asynchronous systems} (which are not synchronous).

		For those systems, the \emph{Earliest Deadline First} (EDF) algorithm, in
		which jobs are prioritized according to their absolute deadline, has been
		shown to be optimal \cite{liu1973scheduling}, which means it
		correctly schedules any feasible system within that group.

@@

	We recall the necessary and sufficient condition explained in \cite{baruah1999generalized, baruah1990algorithms,pellizzoni2005feasibility}, based on the demand-bound function.

@@

	Mathematically,
	\[
		\dbf{t_1, t_2} = \sum_{i=0}^{n-1} n_i(t_1, t_2) \, C_i
	\]
	where $n_i(t_1, t_2)$ is the number of jobs of task $i$ which arrival times
	and deadlines are both in the closed interval $[t_1, t_2]$.

	The values of the $n_i$ are given by
	\[
		n_i(t_1, t_2) =
		\left\llbracket
			\left\lfloor
				\frac{t_2 - O_i - D_i}{T_i}
			\right\rfloor -
			\left\lceil
				\frac{t_1 - O_i}{T_i}
			\right\rceil + 1
		\right\rrbracket_0
	\]
	where $\llbracket x \rrbracket_0$ stands for $\max \{ 0, x \}$.

@@

			The authors of \cite{george2009characterization} give the following definition of the C-space (based on \cite{bini2004schedulability}):
			\begin{definition}
				The \textbf{C-space} of a task system $\tau$ is a region of $n$ dimensions (where each dimension denotes the possible $C_i$ of a task of $\tau$) such that for any vector $C = \{ C_0, \cdots, C_{n-1}\}$ in it, $\tau$ is feasible.
			\end{definition}

@@

      As explained in \cite{nemhauser1988integer}, a region of $\mathbb{N}_0^n$ such as the C-space can be described by a \emph{polytope}, or a list of parametric linear constraints (called the \emph{formulation}). The constraints must be conditions which can unequivocally decide if a point is included in the region or not. In the sequel we give a formulation for the C-space in $\mathbb{N}_0^n$ (For a formulation in $\mathbb{N}^n$, we can add $C_i \geqslant 1 \; \forall i$ to the list of constraints).

      The DBF test gives us a list of inequations on the WCET values of the task set which together are a necessary and sufficient condition of feasibility. They can all be written in the following way:
      \[
        \sum_{i=0}^{n-1} n_i(t_1, t_2) \, C_i \leq t_2 - t_1 \; \forall \: 0 \leq t_1 \leq t_2
      \]
      Note that for a given $t$, we have that the $n_i(t)$ are
      constant (w.r.t. the $C_i$). Those inequations are thus indeed linear.

      However, this description contains an infinite number of constraints, most of
      which are redundant. While this does not harm the precision of the
      test, the smallest description is preferred as it can greatly
      reduce the computation time of practical feasibility tests. In Section~\ref{sct:removeRedundancy} a method to reduce the size of this set of constraints is given.

@@

  In the synchronous case, is has been shown in \cite{baruah1990algorithms} that the constraints corresponding
  to every interval $[0, t]$ (with $t$ a job deadline happening before $H$) are
  necessary and sufficient to test the feasibility of a system.

  Even then, the number of constraints to consider is exponential in the number
  of tasks. While it has been shown that constraints that cover intervals ending
  after the end of the first busy period are also redundant, this value depends on the $C_i$ and is
  therefore not usable for the purposes of this paper. The
  following concept is thus introduced, which is similar to the end of the busy period but does not depend on the execution times.

@@

  The value of the first DIT can be found by solving the following modular equation system:
  \[
    t_d = \min
    \left( t \; s.t. \; \forall i:
      \left\{
        \begin{array}{c}
          t > 0 \\
          t \equiv a_i \; (mod \; T_i) \\
          a_i \in [D_i, T_i]
        \end{array}
      \right.
    \right)
  \]
  The authors of \cite{george2009characterization} give a method to solve this system for cases in which periods are pairwise coprime.

  For synchronous systems in which the execution times are not known, the first DIT $t_d$
  is our ``best guess'' at an upper limit for the end of the busy period. As it was sufficient
  to check the DBF test in intervals $[0, t]$ before the end of the busy period, it is
  slightly more pessimistic but still correct to use the following condition:
  \[
  	\begin{array}{c}
  		\dbf{0,t} \leqslant t \; \forall t \leqslant t_d \\
  		\iff \text{The system is feasible}
  	\end{array}
  \]

@@

  \subsection{Existence condition}
  \label{sct:FPDITexist}

		As in the synchronous case, the problem of existence of the FPDIT
		can be written as a system of modular equations:
		\[
			\begin{array}{l}
				\exists ? \; t_d \; s.t. \; \forall i \; :\\
				\left\{
					\begin{array}{l}
						t_d > O_i \\
						t_d - O_i \equiv a_i \; (mod \; T_i) \\
						a_i \in [D_i, T_i]
					\end{array}
				\right.
			\end{array}
		\]

@@

		\begin{center}
			\begin{tabular}{|r|c|c|c|c|}
				\hline
							& $O_i$ & $C_i$ & $D_i$ & $T_i$ \\ \hline
				$\tau_1$ 	& 0 	& $C_1$ & 3 	& 4\\ \hline
				$\tau_2$ 	& 2 	& $C_2$ & 7 	& 8\\ \hline
			\end{tabular}
		\end{center}

@@

	\subsection{Solving the modular system}
		Finding the FPDIT of a system is similar to finding at least one synchronous
		instant (an instant at which all tasks release a job simultaneously) out of a
		collection of systems with different offsets. The existence conditions can
		directly be used to calculate the value of the FPDIT (if there is one),
		because for a given set of $a_i$ values they are very similar to an instance
		of the Chinese Remainder Problem (CRP), which is easily solved using Gauss's
		CRP solving method for non-coprime $T_i$ \cite{gauss1965disquisitiones}.

		However this is very inefficient as there are exponentially many different
		combinations of $a_i$ values in the number of tasks. A better method for finding the FPDIT consists in splitting Gauss's
		algorithm into two separate parts. Its global structure is simply a sum of
		factors, each term being dependent on which $a_i$ value is chosen. A major speedup can
		be achieved by first computing all possible terms for each task then creating all
		possible DIT values by summing the terms in all possible ways. The number of results is still
		exponential in the number of tasks, but the only operation that is repeated
		exponentially many times is now a sum.

@@

	This simulation aims to quantify how much system have a FPDIT and which factor can influence this quantity.

	We thus checked the existence condition on systems generated by:
	\begin{itemize}
		\item Generating utilization values with the \texttt{UUniFast} algorithm described in \cite{bini2005measuring}, with parameters
		\begin{itemize}
			\item $U_{tot}$ (system utilization) chosen uniformly between 0.25 and 0.75
			\item $n$ (number of tasks) equal to 3
		\end{itemize}
		\item Choosing the $T_i$ uniformly between 5 and 20, which also gives us the $C_i = \lfloor U_i \, T_i \rfloor$
		\item Choosing the $O_i$ with a normal distribution of mean $T_{min}$ and standard deviation $\frac{T_{max} - T_{min}}{2}$
		\item Choosing the $D_i$ uniformly in the interval $[T_i - \frac{T_i - C_i}{CDF}, T_i]$, where the $CDF$ is called \emph{Constrained Deadline Factor}.
	\end{itemize}

	The number of tasks and the constrained Deadline Factor is shown to have a direct influence on the density of FPDIT. Indeed, Fig~\ref{fig:noFPDIT} show the percentage of systems having no FPDIT for different values of CDF.

@@

	We observe that the more we are close to the implicit deadline case, the less systems have FPDIT. This is easily explained by the fact that a lower CDF means more idle instant for each task, and thus more opportunities for DITs to happen. Similarly, as all tasks must be synchronized for a DIT to happen, a higher number of tasks results in a decrease in FPDIT density.

@@

		The description we presented in Section \ref{sct:cspaceDescr} was general, and
		is thus still relevant. Therefore the C-space is still accurately described by
		the following infinite number of constraints:

		\[
			\sum_{i=0}^{n-1} n_i(t_1, t_2)
			\, C_i \leq t_2 - t_1 \; \forall \: 0 \leq t_1 \leq t_2
		\]

@@

		\subsubsection{Job arrival and deadline}
			Consider an interval $[t_1, t_2]$. If $t_2$ is an instant without any job
			deadline, then by definition the value of $\dbf{t_1, t_2}$ will be equal to
			$\dbf{t_1, t_2^*}$, where $t_2^*$ is the latest deadline before $t_2$.

@@

		\subsubsection{Feasibility interval}
			First let us recall how an asynchronous system behaves under an optimal
			scheduler, for example EDF.

@@

As shown in Fig. \ref{fig:asyncBehavior}, the system begins in the \emph{incomplete period}, where all tasks are not in the system. Then, after $O_{max}$ time units, every tasks is in the system and the \emph{transitive period} begins. Then, after $O_{max} + H$ time units, the system is faced with the same pattern of arrival as in $O_{max}$. However, because in $O_{max} + H$ every task is in the system and in $O_{max}$ some were missing, the state of the system will be different. The \emph{stationary period} then begins. It is only after $O_{max} + 2H$ time units that the system comes back into a state in which we can be sure it has been before.

@@

The FPDIT, if it exists, must occur in the transitive period. Indeed, it cannot occur in the incomplete period by definition ($t_d > O_{max}$) and if it occurred in the $k^{th}$ stationary period, another DIT should have occurred in the transitive period at instant $t_d - k H$ (a contradiction).

As the FPDIT happens when all tasks are in the system, we know that the system will arrive in the exact same state one hyperperiod later ($t_d + H$). We can thus restrict the DBF test to intervals included in the interval $[t_d, t_d + H]$, if the FPDIT exists. This will always be a better interval than $[O_{max}, O_{max} + 2H]$ and if the first periodic DIT does not exist (which can be verified as explained in Section \ref{sct:FPDITexist}), the latter interval can be used.

@@

Such a linear problem can be solved by e.g. the simplex algorithm and a branch-and-cut approach. However, in our case the problem is not exactly the same as we have a set of constraints and we want to find all redundant constraints within. A naive method would be to test each constraint against all the others. We present a similar method with an efficient preprocessing when most constraints are redundant in Algorithm \ref{alg:prunRedun}.

@@

After doing this, no redundant constraints are left. However, this algorithm takes an exponential time in the number of constraints. An efficient initial description of $CSPACE$, as given by the methods described in previous sections, is therefore preferred.

@@

For each $[a, d]$, we have a constraint $$\sum_{i=0}^{n-1} n_i(a, d) \, C_i \leqslant d - a$$ which give a description of the C-space.

We now apply Algorithm \ref{alg:prunRedun} to remove redundancies, leaving us with the following constraints:
$$
\left\{
  \begin{array}{ccccc}
    & & C_2 & \leqslant & 2 \\
    C_1 & + & C_2 & \leqslant & 7
  \end{array}
\right.
$$
The C-space thus contains 11 points $(C_1, C_2) = (1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(6, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$, $(4, 2)$ and $(5, 2)$.

@@

    \begin{center}
    \begin{tabular}{|r|c|c|c|c|}
     \hline
      & $O_i$ & $C_i$ & $D_i$ & $T_i$ \\
     \hline
     $\tau_1$ & 0 & $C_1$ & 7 & 15\\
     \hline
     $\tau_2$ & 0 & $C_2$ & 2 & 5\\
     \hline
    \end{tabular}
    \end{center}

The first DIT happens at time $t_d = 7$. The intervals to consider are thus of the form $[0, t]$ where $t$ is a deadline happening before or at time $t_d$. This yields 2 intervals ($[0, 2]$ and $[0, 7]$) and the following constraints describing the C-space, none of them being redundant:
$$
\left\{
  \begin{array}{ccccc}
    & & C_2 & \leqslant & 2 \\
    C_1 & + & 2 C_2 & \leqslant & 7
  \end{array}
\right.
$$
This time, the C-space only contains 8 points: $(1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$.

@@

% conference papers do not normally have an appendix

% use section* for acknowledgement
% \section*{Acknowledgment}

% The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \nocite{*}
\bibliographystyle{plain}
\bibliography{dit-paper}

% that's all folks
\end{document}

>>>>

%
%  $Description: Author guidelines and sample document in LaTeX 2.09$
%
%  $Author: ienne, modified by Laurent George $
%  $Date: 2012/02/02 15:20:59 $
%  $Revision: 1.5 $

\documentclass[times, 10pt,twocolumn, a4paper]{article}
\usepackage{latex8}
\usepackage{epsfig, hhline}
\usepackage{amssymb}
\usepackage{vmargin,array,multirow,float,shadow,times, euscript}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{pdftricks}
\usepackage{stmaryrd}

@@

%\newenvironment{accolade}{\begin{centering} $$ \left\{ \begin{array}{l}}{\end{array} \right. $$ \end{centering}}

%-------------------------------------------------------------------------
% take the % away on next line to produce the final camera-ready version
\pagestyle{empty}

%-------------------------------------------------------------------------
\begin{document}

\title{Describing the C-Space of Asynchronous Periodic Task Systems using
Definitive Idle Times}

\author{Thomas Chapeaux, Paul Rodriguez\\
Universit\'e Libre de Bruxelles\\ tchapeau@ulb.ac.be, paurodri@ulb.ac.be \\
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
\and
Laurent George\\
ECE Paris\\
First line of institution2 address\\ Second line of institution2 address\\
lgeorge@ece.fr\\
}

\maketitle
\thispagestyle{empty}

\begin{abstract}
	Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla lobortis mi vel
turpis ultricies vulputate vel at arcu. Curabitur lectus metus, rutrum sed malesuada ut, mollis sit amet mauris. Integer dapibus in risus et porttitor. In hac habitasse platea dictumst. Maecenas dolor libero, suscipit porta consequat eu, consequat non orci. Sed tortor neque, tincidunt congue cursus sit amet, luctus id ipsum. Aenean tempor est nisl, at dictum metus elementum iaculis.\\

Keywords: Real-time systems, C-space, WCET, sensitivity analysis
\end{abstract}

@@

	Abstract models for real-time systems often require the determination of the
	(worst-case) computation time of a job (WCET), a value highly dependent on the
	platform on which the application will be deployed.\\

	The \emph{C-space} of a real-time system is, intuitively,
	the set of WCET values for which this system would be feasible. A
	precise and efficient description of its C-space would allow to quickly decide
	if a system is feasible on a given platform.\\

@@

		We consider a discrete timeline and tasks systems $\tau$ comprised of periodic concrete
		tasks with constrained deadlines on uniprocessor platforms. Each task
		$\tau_i$ is represented by a tuple $(O_i, C_i, D_i, T_i)$ where $O_i$ is the offset value,
		$C_i$ the WCET, $D_i$ the relative deadline and $T_i$ the period.
		The constrained deadline property implies that $D_i \leq T_i \; \forall i$. In
		the following, we differentiate between \emph{synchronous systems} (where $O_i
		= 0 \; \forall i$) and \emph{asynchronous systems} (which are not synchronous).\\

		For those systems, the \emph{Earliest Deadline First} (EDF) algorithm, in
		which jobs are prioritized according to their absolute deadline, has been
		shown to be optimal \cite{liu1973scheduling}, which means it
		correctly schedules any feasible system within that group.\\

@@

	We recall the necessary and sufficient condition explained in \cite{baruah1999generalized, baruah1990algorithms}, based on the demand-bound function.

@@

	Mathematically,
	\[
		\dbf{t_1, t_2} = \sum_{i=0}^{n-1} n_i(t_1, t_2) \, C_i
	\]
	where $n_i(t_1, t_2)$ is the number of jobs of task $i$ which arrival times
	and deadlines are both in the closed interval $[t_1, t_2]$.\\

	The values of the $n_i$ are given by
	\[
		n_i(t_1, t_2) =
		\left\llbracket
			\left\lfloor
				\frac{t_2 - O_i - D_i}{T_i}
			\right\rfloor -
			\left\lceil
				\frac{t_1 - O_i}{T_i}
			\right\rceil + 1
		\right\rrbracket_0
	\]
	where $\llbracket x \rrbracket_0$ stands for $\max \{ 0, x \}$.\\

@@

			The authors of \cite{george2009characterization} give the following definition of the C-space:
			\begin{definition}
				The \textbf{C-space} of a task system $\tau$ is a region of $n$ dimensions (where each dimension denotes the possible $C_i$ of a task of $\tau$) such that for any vector $C = \{ C_0, \cdots, C_{n-1}\}$ in it, $\tau$ is feasible.
			\end{definition}

%Thus, with a good description of the C-space, it is easy to check if a given vector of $C_i$ value (given e.g. when deploying the application on a specific platform) allows the system to be scheduled or not.

@@

			As explained in \cite{nemhauser1988integer}, a region of $\mathbb{N}_0^n$ such as the C-space can be described by a \emph{polytope}, or a list of parametric linear constraints (called the \emph{formulation}). The constraints must be conditions which can unequivocally decide if a point is included in the region or not. In the sequel we give a formulation for the C-space in $\mathbb{N}_0^n$ (For a formulation in $\mathbb{N}^n$, we can add $C_i \geqslant 1 \; \forall i$ in the list of constraints).\\

			The DBF test gives us a list of inequations on the WCET values of the task set which together are a necessary and sufficient condition of feasibility. They can all be written in the following way:
			\[
				\sum_{i=0}^{n-1} n_i(t_1, t_2) \, C_i \leq t_2 - t_1 \; \forall \: 0 \leq t_1 \leq t_2
			\]
			Note that for a given $t$, we have that the $n_i(t)$ are
			constant (w.r.t. the $C_i$). Those inequations are thus indeed linear.\\

			However, this description contains an infinite number of constraints, most of
			which are redundant. While this does not harm the precision of the
			test, the smallest description is preferred as it can greatly
			reduce the computation time of practical feasibility tests using this system
			of inequations. In Section~\ref{sct:removeRedundancy} a method to reduce the size of this set of constraints is given.

@@

	In the synchronous case, is has been shown that the inequations corresponding
	to every interval $[0, t]$ (with $t$ a job deadline happening before $H$) are
	necessary and sufficient to test the feasibility of a system.\\

	Even then, the number of inequations to consider is exponential in the number
	of tasks. While it has been shown that constraints that cover intervals ending
	after the end of the first busy period are also redundant, this value depends on the $C_i$ and is
	therefore not usable, or at least not for the purposes of this paper. The
	following concept is thus introduced, which is similar to the end of the busy period but does not depend on the execution times.

@@

	The value of the first DIT can be found by solving the following modular equation system:
	\[
		t_d = \min
		\left( t \; s.t. \; \forall i:
			\left\{
				\begin{array}{c}
					t > 0 \\
					t \equiv a_i \; (mod \; T_i) \\
					a_i \in [D_i, T_i]
				\end{array}
			\right.
		\right)
	\]
	The authors of \cite{george2009characterization} give a method to solve this system for cases in which periods are pairwise coprime.\\

	Every DIT being an idle instant, the execution of the system after a DIT does
	not depend on the execution before it. Because the synchronous arrival pattern is
	the worst case [citation needed], it is sufficient for the DBF test to only
	consider intervals $[0, t]$ with $t$ being a deadline before or at the first DIT.\\

@@

	\subsection{Existence condition}
  \label{sct:FPDITexist}
		As in the synchronous case, the problem of existence of the FPDIT
		can be written as a system of modular equations:
		\[
			\begin{array}{l}
				\exists ? \; t_d \; s.t. \; \forall i \; :\\
				\left\{
					\begin{array}{l}
						t_d > O_i \\
						t_d - O_i \equiv a_i \; (mod \; T_i) \\
						a_i \in [D_i, T_i]
					\end{array}
				\right.
			\end{array}
		\]

@@

		\begin{center}
		\begin{tabular}{|r|c|c|c|c|}
		 \hline
		  & $O_i$ & $C_i$ & $D_i$ & $T_i$ \\
		 \hline
		 $\tau_1$ & 0 & $C_1$ & 3 & 4\\
		 \hline
		 $\tau_2$ & 2 & $C_2$ & 7 & 8\\
		 \hline
		\end{tabular}
		\end{center}
		~\\

@@

	\subsection{Solving the modular system}
		Finding the FPDIT of a system is similar to finding a synchronous instant of
		this system (an instant at which all tasks release a job simultaneously), but
		with weaker constraints. That is, the FPDIT is an instant after the last
		deadline of all tasks, but before the next arrival time of all tasks. Finding
		a synchronous instant is equivalent to finding a solution $t$ for the
		following set of constraints :
		\[
			\left\{
			\begin{array}{l}
				t \geqslant O_i \: \forall \: \tau_i \in \tau \\
				t \equiv O_i \; (mod \; T_i) \: \forall \: \tau_i \in \tau \\
			\end{array}
			\right.
		\]
		which is very similar to an instance of the Chinese Remainder Problem (CRP). It
		is not possible to directly use a solver for the CRP (such as Gauss's
		algorithm \cite{gauss1965disquisitiones}) as in a task system, $T_i$ values could have common factors.
		However, the instance can be modified (each constraint with a non-prime period can be split into powers of primes
		without changing the solution if there is one) in order to make the problem
		solvable through applying a solution for the CRP. \\

		Now that we have a solver for the problem of finding a synchronous instant, we
		actually also have a solver for finding the FPDIT. The problem of finding the
		first DIT is the same as finding all the possible DIT then taking the minimum.
		Each of those DIT corresponds to a unique combination of $a_i$ values (one
		for each task). Finding one of such DIT is the same problem as finding a
		synchronous instant :
		\[
			\left\{
			\begin{array}{l}
				t \geqslant O_i + a_i \: \forall \: \tau_i \in \tau \\
				t \equiv O_i \; (mod \; T_i) \: \forall \: \tau_i \in \tau \\
			\end{array}
			\right.
		\]
		all DITs can be found by solving the equivalent CRP problem for all
		combinations of $a_i$ values. This is however very inefficient as there are
		exponentially many different combinations of $a_i$ values in the number of
		tasks.\\

		A better method for finding the FPDIT consists in splitting Gauss's algorithm
		into two separate parts. The global structure is simply a sum of factors, each
		term being dependent on which $a_i$ value is chosen. A major speedup can
		be achieved by first computing all possible terms for each task (only one
		modular inverse must be computed per task that way) then creating all possible DIT values
		by summing the terms in all possible ways. The number of results is still
		exponential in the number of tasks, but the only operation that is repeated
		exponentially many times is now a sum.

@@

	We wanted to quantify how much systems have a FPDIT........

	We generated task systems by
	\begin{itemize}
		\item Generating utilization values with the \texttt{UUniFast} algorithm described in \cite{bini2005measuring}, with parameters
		\begin{itemize}
			\item $U_{tot}$ (system utilization) chosen uniformly between 0.25 and 0.75
			\item $n$ (number of tasks) equal to 3
		\end{itemize}
		\item Choosing the $T_i$ uniformly between 5 and 20, which also gives us the $C_i = \lfloor U_i \, T_i \rfloor$
		\item Choosing the $O_i$ with a normal distribution of mean $T_{min}$ and standard deviation $\frac{T_{max} - T_{min}}{2}$
		\item Choosing the $D_i$ uniformly in the interval $[T_i - \frac{T_i - C_i}{CDF}, T_i]$, where the $CDF$ is called \emph{Constrained Deadline Factor}.
	\end{itemize}
	and then computed the value of the FPDIT.\\

	We use the CDF because........\\

@@

	The result are represented in Fig. \ref{fig:noFPDIT}. We observe that the more we are close to the implicit deadline case, the less systems have FPDIT.

@@

  The description we presented in Section \ref{sct:cspaceDescr} was general, and is thus still relevant. Therefore the C-space is still accurately described by the following infinite number of constraints:$$\sum_{i=0}^{n-1} n_i(t_1, t_2) \, C_i \leq t_2 - t_1 \; \forall \: 0 \leq t_1 \leq t_2$$

@@

\subsubsection{Job arrival and deadline}

Consider an interval $[t_1, t_2]$. If $t_2$ is an instant without any job deadline, then by definition the value of $\dbf{t_1, t_2}$ will be equal to $\dbf{t_1, t_2^*}$, where $t_2^*$ is the latest deadline before $t_2$.\\

@@

  \subsubsection{Feasibility interval}

  First let us recall how an asynchronous system behaves under an optimal scheduler, for example EDF.\\

@@

As shown in Fig. \ref{fig:asyncBehavior}, the system begins in the \emph{incomplete period}, where all tasks are not in the system. Then, after $O_{max}$ time units, every tasks is in the system and the \emph{transitive period} begins. Then, after $O_{max} + H$ time units, the system is faced with the same pattern of arrival as in $O_{max}$. However, because in $O_{max} + H$ every task is in the system and in $O_{max}$ some were missing, the state of the system will be different. The \emph{stationary period} then begins. It is only after $O_{max} + 2H$ time units that the system comes back into a state in which we can be sure it has been before.\\

@@

The FPDIT, if it exists, must occur in the transitive period. Indeed, it cannot occur in the incomplete period by definition ($t_d > O_{max}$) and if it occurred in the $k^{th}$ stationary period, another DIT should have occurred in the transitive period at instant $t_d - k H$ (a contradiction).\\

As the FPDIT happens when all tasks are in the system, we know that the system will arrive in the exact same state one hyperperiod later ($t_d + H$). We can thus restrict the DBF test to intervals included in the interval $[t_d, t_d + H]$, if the FPDIT exists. This will always be a better interval than $[O_{max}, O_{max} + 2H]$ and if the first periodic DIT does not exist (which can be verified as explained in Section \ref{sct:FPDITexist}), the latter interval can be used.\\

@@

Such a linear problem can be solved by e.g. the simplex algorithm and a branch-and-cut approach. However, in our case the problem is not exactly the same as we have a set of constraints and we want to find all redundant constraints within. A naive method would be to test each constraint against all the others. We present a similar method with an efficient preprocessing when most constraints are redundant in Algorithm \ref{alg:prunRedun}.\\

@@

After doing this, no redundant constraints are left. However, this algorithm takes an exponential time in the number of constraints. An efficient initial description of $CSPACE$, as given by the methods described in previous sections, is therefore preferred.\\

@@

For each $[a, d]$, we have a constraint $$\sum_{i=0}^{n-1} n_i(a, d) \, C_i \leqslant d - a$$ which give a description of the C-space.\\

We now apply Algorithm \ref{alg:prunRedun} to remove redundancies, leaving us with the following constraints:
$$
\left\{
	\begin{array}{ccccc}
		& & C_2 & \leqslant & 2 \\
		C_1 & + & C_2 & \leqslant & 7
	\end{array}
\right.
$$
The C-space thus contains 11 points $(C_1, C_2) = (1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(6, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$, $(4, 2)$ and $(5, 2)$.\\

@@

		\begin{center}
		\begin{tabular}{|r|c|c|c|c|}
		 \hline
		  & $O_i$ & $C_i$ & $D_i$ & $T_i$ \\
		 \hline
		 $\tau_1$ & 0 & $C_1$ & 7 & 15\\
		 \hline
		 $\tau_2$ & 0 & $C_2$ & 2 & 5\\
		 \hline
		\end{tabular}
		\end{center}
		~\\

The first DIT happens at time $t_d = 7$. The intervals to consider are thus of the form $[0, t]$ where $t$ is a deadline happening before or at time $t_d$. This yields 2 intervals ($[0, 2]$ and $[0, 7]$) and the following constraints describing the C-space, none of them being redundant:
$$
\left\{
	\begin{array}{ccccc}
		& & C_2 & \leqslant & 2 \\
		C_1 & + & 2 C_2 & \leqslant & 7
	\end{array}
\right.
$$
This time, the C-space only contains 8 points: $(1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$.\\

@@

% \nocite{*}
\bibliographystyle{latex8}
\bibliography{dit-paper}
\end{document}
